---
title: "`Project/notebooks/data_cleaning.R`"
author: "TECK Squad"
date: "December 12, 2025"
output:
  rmdformats::readthedown:
    toc_depth: 2        
    code_folding: hide
fontsize: 11pt
geometry: margin=1in
editor_options:
  chunk_output_type: console
---
  
```{r setup, include=FALSE}
# ---- Setup ----
suppressPackageStartupMessages({
  library(tidyverse)
  library(arrow)
  library(dplyr)
  #library(rprojroot)
  library(knitr)
  library(kableExtra)
  
  #library(broom)
  #library(car)
  #library(latex2exp)
  #library(MPV)
  #library(glmnet)
})

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 6.5,
  fig.height = 4
)
theme_set(theme_bw(base_size = 12))

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

# ---- Helper functions ----
fmt   <- function(x, d = 3)
  round(x, d)
fmt_e <-
  function(x, digits = 3)
    formatC(x, format = "e", digits = digits)

# ---- Load data I/O utilities and raw data ----
source("Project/scripts/data_io.R")
# Raw data from parquet via `data_io.R`
ad_raw <- load_ad_data(preview = FALSE)

# ---- Data Log setup ----
cleaning_log <- tibble::tibble(
  step_id = character(),
  variable = character(),
  issue = character(),
  decision = character(),
  rows_affected = integer()
)
```

---

# **Purpose**

This notebook documents the full data-cleaning workflow for the Oregon ad bidding dataset.

  - **Input:** raw parquet data loaded via `load_ad_data()` (see `Project/scripts/data_io.R`).
  - **Output:** a cleaned analysis-ready dataset saved under `Project/data/...`.

---

---

# **DATA OVERVIEW**

- According to the data dictionary, each row represents **one bidder’s response for one requested size within a single ad auction**, meaning:
  - Each `AUCTION_ID` appears multiple times  
  - There may be several responses for the same ad size within the same auction  

---

## Dimensions and Structure

We begin with a high-level look at the raw dataset.

```{r overview-dims, message=FALSE, 'hold'=TRUE}
# Dimensions
n_rows <- nrow(ad_raw)
n_cols <- ncol(ad_raw)

dim_table <- tibble(
  Metric = c("Number of Rows", "Number of Columns"),
             Value = c(n_rows, n_cols)
)

dim_table %>%
  kable(caption = "Dataset Dimentions") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped"))
```

---

## Basic summaries

We summarize all variables numerically and inspect missingness patterns using the helper summary function defined in `data_io.R`.

```{r overview-summary}
summarize_ad_data(ad_raw)
```

---

## Expected vs. Actual Data Types

We can compare expected data types as given by the data dictionary to the actual type of each variable in the data:

```{r overview-expected-vs-actual, message=FALSE}
expected_types <- tibble(
  Variable = c(
    "TIMESTAMP",
    "DATE_UTC",
    "AUCTION_ID",
    "PUBLISHER_ID",
    "DEVICE_TYPE",
    "DEVICE_GEO_COUNTRY",
    "DEVICE_GEO_REGION",
    "DEVICE_GEO_CITY",
    "DEVICE_GEO_ZIP",
    "DEVICE_GEO_LAT",
    "DEVICE_GEO_LONG",
    "REQUESTED_SIZES",
    "SIZE",
    "PRICE",
    "RESPONSE_TIME",
    "BID_WON"
  ), 
  Expected_Class = c(
    "datetime",          # TIMESTAMP_NTZ
    "date",              # DATE
    "categorical",       # Auction ID
    "categorical",       # Publisher ID
    "categorical",       # desktop/mobile/tablet
    "categorical",       # 2-letter country
    "categorical",       # 2-letter USPS region
    "categorical",       # City name
    "categorical",       # ZIP / ZIP+4 string
    "numeric",           # Latitude
    "numeric",           # Longitude
    "categorical",       # List of sizes
    "categorical",       # Single size
    "numeric",           # CPM bid
    "numeric",           # Response time in ms
    "boolean"            # TRUE/FALSE winner
  ),
  Expected_Detail = c(
    "TIMESTAMP_NTZ (UTC event time)",
    "DATE (UTC calendar date)",
    "VARCHAR (auction identifier)",
    "VARCHAR (publisher/site identifier)",
    "VARCHAR, (device category, string)",
    "VARCHAR(2) (ISO 2-letter country code)",
    "VARCHAR(2) (USPS state/region postal code)",
    "VARCHAR, (city name, string)",
    "VARCHAR(10) (ZIP or ZIP+4)",
    "FLOAT (latitude of device)",
    "FLOAT (longitude of device)",
    "VARCHAR/ARRAY (all requested sizes, ex: '300x250, 320x50')",
    "VARCHAR (size for this bid, ex: `300x250')",
    "NUMBER(12,6) (CPM bid in USD)",
    "NUMBER(10,0) (bidder latency in ms)",
    "BOOLEAN (if bid won, TRUE/FALSE)"
  )
)

expected_types <- expected_types %>%
  arrange(factor(
    Expected_Class,
    levels = c("numeric", "categorical", "datetime", "date", "boolean")
  ))

actual_types <- tibble(
  Variable = names(ad_raw),
  Actual_Type = sapply(ad_raw, class)
)

comparison <- expected_types %>%
  left_join(actual_types, by = "Variable")

comparison %>%
  kable(caption = "Expected vs Actual Types (Data Dictionary vs Raw Data)") %>%
    kable_styling(full_width = FALSE, bootstrap_options = c("striped"))
```

Many of these are suspicious or inconsistent with the data dictionary:
  
  - **`PRICE`** and **`RESPONSE_TIME`** 
      - Expected *numeric* $\to$ Stored as `character`
      - Likely due to *non-numeric characters* or inconsistent formatting.
  - **`DEVICE_TYPE`**
      - Expected descriptive *string* $\to$ Stored as an `integer` code
      - Implies an encoded representation without a visible lookup table.
  - **`DEVICE_GEO_COUNTRY`** 
      - Expected per dictionary $\to$ Not present in this data subset
  - **`TIMESTAMP`** and **`DATE_UTC`** 
      - Expected *datetime/date* $\to$ Stored as `character`
      - Will need to be converted prior to any time-based analysis.
  - **`BID_WON`** 
      - Expected *boolean* $to$ Stored as `character`
      - Will require re-coding to `logical` (TRUE/FALSE) type.

In addition, some geographic and categorical fields (**`DEVICE_GEO_REGION`**, **`DEVICE_GEO_CITY`**, and **`DEVICE_GEO_ZIP`**) appear broadly consistent with expected classes, but will need to be fully analysed for any messy real-world values such as **inconsistent casing**, **invalid state codes**, or **irregular ZIP formats**. These issues will be fully examined and addressed in the variable-level cleaning sections that follow.

---

---

# **FULL DATASET**

---

## Duplicated rows

We check for exact duplicate rows in `ad_raw` and drop them if present.

- Check for exact duplicate rows
- Eliminate & update log

---

## Keys

We confirm the expected grain and check uniqueness of key fields (e.g., `AUCTION_ID` and combinations with bidder/seat/size, as appropriate for this subset).

- Check key uniqueness (e.g., `AUCTION_ID` + other IDs if relevant)

---

---

# **VARIABLE**
   
## Description

  - Description & expected type/constraints (from dictionary).
   - Use the dictionary to define expectations.
   
## Exploration

  - values, distributions, weird codes, NA patterns.

## Cleaning

   -type conversion, recodes, outlier handling, bounds checks, etc.

## Summary

   -Add to cleaning_log if we changed anything.
   
---

---

# **CLEANED DATASET**

## Assemble `ad_clean`

## # Post-Cleaning Inspection

  - Final sanity checks on NA structure, bounds, duplicates

## Save via `data_io.R` function `export_ad_data()`

---

---

# **DATA LOG**

  - Display cleaning_log table.
  
  
  
  
  
---